---
title: "1. Macrofauna Alpha Diversity"
description: |
  Reproducible workflow to assess the macrofaunal diversity. 
author:
#  - name: Noelle Lucey
#    url: https://example.com/norajones
#    affiliation: Spacely Sprockets
#    affiliation_nrl: https://example.com/spacelysprokets
bibliography: assets/cite.bib
output:
    distill::distill_article:
      css: assets/styles.css
      toc: true
      toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE) # echo=true, visible code, eval=true, with output

library(vegan)
library(ggplot2)
library(cowplot)
library(mvabund) 
library(ggpubr)
library(scales)
```

> Hit the *Hide Code* button to hide the R code.

<aside>

```{r codefolder_ssu, echo=FALSE, results='asis', eval=TRUE}
codefolder::generic(init = "show", query = "pre.sourceCode",
  style = "position: absolute; right: 14%; z-index: 200")
```

</aside>

# Summary

This workflow contains diversity assessments for the **full** data set---all samples---and a condensed **short** data set---samples averaged by site and season. In order to run the workflow, you need the output files T_REPO_raw.csv or macro_short1.csv. See the [Data Availability](data-availability.html) page for complete details.


```{r, include=FALSE}
ds <- read.csv("files/trepo/macrodata/T_REPO_raw.csv", header = TRUE)
ds <- ds[ds$size == "combined",] # select combinded size fraction
ds$date <- as.Date(ds$date, format = "%m/%d/%y", tz = "EST")
ds$site <- ordered(ds$site, levels = c("Pastores", "Almirante", "Cristobal", "P_Caracol"))
ds <- ds[ds$week != "WXX",] # only looking at one year period corresponding to the microbial dataset

ds.data <- ds[,16:32] # species matrix
env.data <- ds[,34:40] # temp, DO %, DO mgL, pH, chloro, bga
site.data <- ds[,1:7] # site, region, season, week, Region, season
str(site.data)
# Data Exploration with truncated data set, from dates matching microbial samples
# This is the sum of occurrences at each site, in each season every other week (based on microbial sampling); environmental data is averaged by season and site
short <- read.csv("files/trepo/macrodata/macro_short1.csv", header = TRUE)
short$site <- ordered(short$site, levels = c("Pastores", "Almirante", "Cristobal", "P_Caracol"))
short.data <- short[,13:29] # species
env_s.data <- short[,31:39] # temp, DO %, DO mgL, pH, chloro, bga
site_s.data <- short[,1:3] # site, region, season
```

```{r, include=FALSE}
swel_col <- c("#E69F00", "#CC79A7", "#0072B2", "#56B4E9")
```

The macrofaunal community; raw untransformed counts:

```{r}
par(mar=c(2,10,2,2)) # adjusts the margins
boxplot(ds[,16:32],horizontal = TRUE,las=2, main="Abundance")
```

# Calculate Alpha diversity

    A)  Shannon's Diversity H'
    B)  Observed Richness (\# of species)
    C)  Pielou's Evennness
    D)  Inverse Simposon Diversity
    E)  Renyi's Entropy

```{r}
# Shannon's H'
H_short <- diversity(short.data)
H_long <- diversity(ds.data)

# Observed Richness
richness_short <- specnumber(short.data)  
richness_long <- specnumber(ds.data) 

# Pielou's Evenness
evenness_short <- H_short/log(richness_short)
evenness_long <- H_long/log(richness_long)

# Inverse simpson
inv_simpson_long <- diversity(ds.data, index = "invsimpson", base = 2)
inv_simpson_short <- diversity(short.data, index = "invsimpson", base = 2)

# Renyi entropy as diversity measure
R <- renyi(ds.data, scales = 2)
N2 <- renyi(ds.data, scales = 2, hill = TRUE)  # inverse simpson
k <- sample(nrow(ds.data), 6)
R <- renyi(ds.data[k,])

Rs <- renyi(short.data, scales = 2)
N2s <- renyi(short.data, scales = 2, hill = TRUE)  # inverse simpson
ks <- sample(nrow(short.data), 8)
Rs <- renyi(short.data[ks,])
```

Some data exploration... look at the Renyi entropy of the 8 samples from the condensed data-set. This includes a plot for each site and season, and it is clear there are differences between the sites and seasons. There is also some evidence of differences through 'time' (i.e. x-axis) at some of the sites/seasons. We are looking at where the circles fall within the mean (pink dashed lines) and standard deviation (green dashed lines) in each sample. 

```{r}
plot(Rs, main = "Renyi Diversity plot; shortened dataset")
```

Create alpha diversity data-frame including environmental data.

```{r}
# long data 
alpha <- cbind(shannon = H_long, richness = richness_long, pielou = evenness_long, inv_simpson = inv_simpson_long, env.data, site.data)

# short data
alpha_s <- cbind(shannon = H_short, richness = richness_short, pielou = evenness_short, inv_simpson = inv_simpson_short, env_s.data, site_s.data)
```

# Plot Alpha Diversity

These are violin plots to visualize the distribution of the diversity
metrics at each site.

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
plot.shan <- ggplot(alpha, aes(x = site, y = shannon, fill = site)) +
  geom_violin() +
  scale_fill_manual(values = swel_col) +
  ylab("Shannon's H'") + 
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4)) +
  stat_compare_means(method = "anova", label.y = 2.5)+      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")

plot.rich <-ggplot(alpha, aes(x = site, y = richness, fill = site)) +
  geom_violin() +
  scale_fill_manual(values = swel_col) +
  ylab("Species Richness") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4)) +
  stat_compare_means(method = "anova", label.y = 13)+      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")

plot.even <- ggplot(alpha, aes(x = site, y = pielou, fill = site)) +
  geom_violin() +
  scale_fill_manual(values = swel_col) +
  ylab("Pielou's Evenness") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4)) +
  stat_compare_means(method = "anova",, label.y = 1.2) +      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")

plot.inv_simpson <- ggplot(alpha, aes(x = site, y = inv_simpson, fill = site)) +
  geom_violin() +
  scale_fill_manual(values = swel_col) +
  ylab("Inverse Simpson") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4)) +
  stat_compare_means(method = "anova", label.y = 8)+      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")
legend <- get_legend(plot.inv_simpson)

plot_grid(plot.shan + theme(legend.position = "none"), 
          plot.rich + theme(legend.position = "none"), 
          plot.even + theme(legend.position = "none"), 
          plot.inv_simpson + theme(legend.position = "none"), ncol = 4)

```

A bit easier to visualize with the condensed, short dataset:

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
plot.shan <- ggplot(alpha_s, aes(x = site, y = shannon, colour = site, shape = season)) +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  ylab("Shannon's H'") + 
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))

plot.rich <-ggplot(alpha_s, aes(x = site, y = richness, colour = site, shape = season)) +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  ylab("Species Richness") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))

plot.even <- ggplot(alpha_s, aes(x = site, y = pielou, colour = site, shape = season)) +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  ylab("Pielou's Evenness") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))

plot.inv_simpson <- ggplot(alpha_s, aes(x = site, y = inv_simpson, colour = site, shape = season)) +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  ylab("Inverse Simpson") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4))
legend <- get_legend(plot.inv_simpson)

plot_grid(plot.shan + theme(legend.position = "none"), 
          plot.rich + theme(legend.position = "none"), 
          plot.even + theme(legend.position = "none"),
          plot.inv_simpson + theme(legend.position = "none"),
          legend, rel_heights = c(1,1,1,1,0.5), ncol = 5)

```

These graphs illustrate the differences between alpha diversity metrics and the relationship between species richness, Pielou's evenness, and Shannon's H'. Punta Caracol and Pastores are two sites with similar Shannon's H' scores, however, this is resultant of distinct mechanisms. P. Caracol has a large number of species (high species richness) but is relatively uneven (low Pielou's Evenness), resulting in a moderate Shannon's H'. Pastores, in contrast, has a small number of species but is very even, also resulting in a moderate Shannon's H'. Alpha diversity is structured very differently at P. Caracol compared with Pastores, but if we had used Shannon's H' alone we may not have identified these important differences.

Almirante has relatively how Shannon's H', which is matched by low species richness and high evenness. From the large difference between the Cristobal point, it appears that the season is important in driving alpha diversity measures (H', richness and evenness to a lesser extent).

# Environmental Effects

Simple linear models between diversity metrics and environmental
conditions.

    A)  Temperature (Celsius)
    B)  Salinity (psu)
    C)  Dissolved Oxygen or 'DO' (mg/L)
    D)  pH (NBS scale)
    E)  Chlorophyll concentration (ug/L)
    F)  Cyanobacteria/Blue Green Algae concentration (ug/L)

First, with the full dataset, assess relationships between Shannon's Diversity and the environmental conditions. Additionally, look at the same relationships within the hypoxic and normoxic season (HS and NS).

```{r, eval=FALSE, echo=TRUE}
summary(lm(shannon ~ temp, alpha))
summary(lm(shannon ~ sal_psu, alpha)) # negative, significant *
summary(lm(shannon ~ DO_mgL, alpha)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha)) # positive, sig. ***
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(shannon ~ temp, alpha1))
summary(lm(shannon ~ sal_psu, alpha1)) 
summary(lm(shannon ~ DO_mgL, alpha1)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha1))  # positive, sig. **
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha1))  
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha1)) # negative, sig * only in reduced data-set

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(shannon ~ temp, alpha2))
summary(lm(shannon ~ sal_psu, alpha2)) # negative, sig. ** 
summary(lm(shannon ~ DO_mgL, alpha2)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha2)) # positive, sig. ***
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha2)) 
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha2)) 
```

<details>

<summary>

Detailed results of Alpha Diversity & Environmental Parameter ANOVA tests

</summary>

```{r, eval=TRUE, echo=TRUE}
summary(lm(shannon ~ temp, alpha))
summary(lm(shannon ~ sal_psu, alpha)) # negative, significant *
summary(lm(shannon ~ DO_mgL, alpha)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha)) # positive, sig. ***
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(shannon ~ temp, alpha1))
summary(lm(shannon ~ sal_psu, alpha1)) 
summary(lm(shannon ~ DO_mgL, alpha1)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha1))  # positive, sig. **
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha1))  
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha1)) # negative, sig * only in reduced data-set

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(shannon ~ temp, alpha2))
summary(lm(shannon ~ sal_psu, alpha2)) # negative, sig. ** 
summary(lm(shannon ~ DO_mgL, alpha2)) # positive, sig. ***
summary(lm(shannon ~ pH, alpha2)) # positive, sig. ***
summary(lm(shannon ~ as.numeric(Chlorophyll_ugL), alpha2)) 
summary(lm(shannon ~ as.numeric(BGA_ugL), alpha2)) 
```

</details>

The linear model outputs demonstrate that there is a significant relationship between DO and pH and Shannon's H' throughout Almirante Bay (in the full dataset). DO and pH are always important regardless of the season. However, in the hypoxic season, lower diversity is related to higher salinity values. This relationship is not present during the normoxic season.

Linear model output demonstrates a significant positive relationship between both DO and pH and Shannon's H' Diversity across Almirante Bay. We can then fit the significant regression lines to our plots using the geom_smooth function in the r package ggplot2.

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
DO.H <- ggplot(alpha_s, aes(x = DO_mgL, y = shannon, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  #geom_smooth(method = "lm", aes(color = season),  fill = "grey90") +
  geom_point(aes(shape = season), size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(DO ~ (mg~L^-1))) + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

DO.H.full <- ggplot(alpha, aes(x = DO_mgL, y = shannon, colour = site, shape=season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = season), size = 2) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(DO ~ (mg~L^-1))) + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

plot_grid(DO.H + theme(legend.position = 'none'),
          DO.H.full + theme(legend.position = 'none'),
          legend, rel_heights = c(1,1,0.5),  ncol = 3)
```

The other significant parameters, with the condensed short dataset:

```{r, echo=FALSE, eval= FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
# plots with full dataset
sal.H <- ggplot(alpha, aes(x = sal_psu, y = shannon, colour = site, shape = season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 2, alpha=0.5) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(salinity ~ (psu))) + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

pH.H <- ggplot(alpha, aes(x = pH, y = shannon, colour = site, shape = season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey85") +
  geom_point(aes(shape = factor(season)), size = 2, alpha=0.5) +
  scale_colour_manual(values = swel_col) +
  xlab("pH") + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

#bga.H <- ggplot(alpha, aes(x = as.numeric(BGA_ugL), y = shannon, colour = site, shape = season)) +
#  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
#  geom_point(aes(shape = factor(season)), size = 2, alpha=0.5) +
#  scale_colour_manual(values = swel_col) +
#  xlab("BGA (ugL)") + 
#  ylab("") +
#  ylim(1,2.3) +
#  theme_bw()
legend <- get_legend(pH.H)

plot_grid(pH.H + theme(legend.position = 'none'),
          sal.H + theme(legend.position = 'none'), 
          #bga.H + theme(legend.position = 'none'),
          legend, rel_heights = c(1,1,0.5),  ncol = 4)
```

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
pH.H <- ggplot(alpha_s, aes(x = pH, y = shannon, colour = site, shape = season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 3, alpha=1) +
  scale_colour_manual(values = swel_col) +
  xlab("pH") + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

sal.H <- ggplot(alpha_s, aes(x = sal_psu, y = shannon, colour = site, shape = season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 3, alpha=1) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(salinity ~ (psu))) + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()

bga.H <- ggplot(alpha_s, aes(x = as.numeric(BGA_ugL), y = shannon, colour = site, shape = season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 3, alpha=1) +
  scale_colour_manual(values = swel_col) +
  xlab("BGA (ugL)") + 
  ylab("") +
  ylim(1,2.3) +
  theme_bw()
legend <- get_legend(bga.H)

plot_grid(pH.H + theme(legend.position = 'none'),
          sal.H + theme(legend.position = 'none'), 
          bga.H + theme(legend.position = 'none'),
          legend, rel_heights = c(1,1,1,0.5),  ncol = 5)
```

Now we look at the same relationships with respect to species richness.

```{r, eval=FALSE, echo=TRUE}
summary(lm(richness ~ temp, alpha))
summary(lm(richness ~ sal_psu, alpha)) # negative, sig.* higher richness with more rain
summary(lm(richness ~ DO_mgL, alpha)) # positive, sig.*** - higher richness with more oxygen
summary(lm(richness ~ pH, alpha)) # positive, sig.*** - higher richness with less acidity
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(richness ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(richness ~ temp, alpha1)) 
summary(lm(richness ~ sal_psu, alpha1)) 
summary(lm(richness ~ DO_mgL, alpha1)) # positive, sig.***
summary(lm(richness ~ pH, alpha1)) 
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha1)) # negative, sig.* - less richness with more chlorophyl
summary(lm(richness ~ as.numeric(BGA_ugL), alpha1)) # negative, sig.* - less richness with more BGA

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(richness ~ temp, alpha2))
summary(lm(richness ~ sal_psu, alpha2)) # negative, sig.*** - less richness with higher salinity 
summary(lm(richness ~ DO_mgL, alpha2)) # positive, sig.*** - higher richness with more oxgyen
summary(lm(richness ~ pH, alpha2)) # positive, sig.*** - higher richness with higher pH
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha2))
summary(lm(richness ~ as.numeric(BGA_ugL), alpha2)) 
```


<details markdown="1">
<summary>Detailed results Richness & Environmental Parameter ANOVA tests</summary>

```{r, eval=TRUE, echo=TRUE}
summary(lm(richness ~ temp, alpha))
summary(lm(richness ~ sal_psu, alpha)) # negative, sig.* higher richness with more rain
summary(lm(richness ~ DO_mgL, alpha)) # positive, sig.*** - higher richness with more oxygen
summary(lm(richness ~ pH, alpha)) # positive, sig.*** - higher richness with less acidity
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(richness ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(richness ~ temp, alpha1)) 
summary(lm(richness ~ sal_psu, alpha1)) 
summary(lm(richness ~ DO_mgL, alpha1)) # positive, sig.***
summary(lm(richness ~ pH, alpha1)) 
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha1)) # negative, sig.* - less richness with more chlorophyl
summary(lm(richness ~ as.numeric(BGA_ugL), alpha1)) # negative, sig.* - less richness with more BGA

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(richness ~ temp, alpha2))
summary(lm(richness ~ sal_psu, alpha2)) # negative, sig.*** - less richness with higher salinity 
summary(lm(richness ~ DO_mgL, alpha2)) # positive, sig.*** - higher richness with more oxgyen
summary(lm(richness ~ pH, alpha2)) # positive, sig.*** - higher richness with higher pH
summary(lm(richness ~ as.numeric(Chlorophyll_ugL), alpha2))
summary(lm(richness ~ as.numeric(BGA_ugL), alpha2)) 
```

</details>

Salinity, oxygen and pH have the greatest effect on species richness. During the normoxic season, increasing chlorophyll and BGA levels correspond to decreasing richness. This is not true in the hypoxic season (no relationship).

Here we can see these relationships by plotting species richness against pH, DO, saliniy, BGA, and chlorophyll.

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
DO.rich <- ggplot(alpha_s, aes(x = DO_mgL, y = richness, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(DO ~ (mg~L^-1))) + 
  ylab("Species Richness") +
  ylim(5,15) +
  theme_bw()

sal.rich <- ggplot(alpha_s, aes(x = sal_psu, y = richness, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(salinity ~ (psu))) + 
  ylab("") +
  ylim(5,15) +
  theme_bw()

pH.rich <- ggplot(alpha_s, aes(x = pH, y = richness, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey85") +
  geom_point(size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab("pH") + 
  ylab("") +
  ylim(5,15) +
  theme_bw()

chloro.rich <- ggplot(alpha_s, aes(x = as.numeric(Chlorophyll_ugL), y = richness, colour = site, shape=season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab("Chlorophyll (ug/L)") + 
  ylab("") +
  ylim(5,15) +
  theme_bw()

bga.rich <- ggplot(alpha_s, aes(x = as.numeric(BGA_ugL), y = richness, colour = site, shape=season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab("BGA (ug/L)") + 
  ylab("") +
  ylim(5,15) +
  theme_bw()
legend <- get_legend(bga.rich)

plot_grid(DO.rich + theme(legend.position = 'none'),
          pH.rich + theme(legend.position = 'none'), 
          sal.rich + theme(legend.position = 'none'), 
          chloro.rich + theme(legend.position = 'none'), 
          bga.rich + theme(legend.position = 'none'), 
          legend, rel_heights = c(1,1,1,1,1,0.5),  ncol = 6)
```

Check the evenness and environmental parameters:

```{r, eval=FALSE, echo=TRUE}
summary(lm(pielou ~ temp, alpha))
summary(lm(pielou ~ sal_psu, alpha)) # positive, significant - higher evenness with more saline
summary(lm(pielou ~ DO_mgL, alpha)) # negative, significant - lower evenness with more oxygen
summary(lm(pielou ~ pH, alpha)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(pielou ~ temp, alpha1)) 
summary(lm(pielou ~ sal_psu, alpha1)) # positive, significant - higher evenness with higher salinity
summary(lm(pielou ~ DO_mgL, alpha1)) 
summary(lm(pielou ~ pH, alpha1)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha1)) 
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha1)) 

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(pielou ~ temp, alpha2))
summary(lm(pielou ~ sal_psu, alpha2)) # marginally significant
summary(lm(pielou ~ DO_mgL, alpha2)) # marginally significant
summary(lm(pielou ~ pH, alpha2)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha2)) # marginally significant
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha2)) # negative, significant - low evenness with more BGA
```


<details markdown="1">
<summary>Detailed results Evenness & Environmental Parameter ANOVA tests </summary>

```{r, eval=TRUE, echo=TRUE}
summary(lm(pielou ~ temp, alpha))
summary(lm(pielou ~ sal_psu, alpha)) # positive, significant - higher evenness with more saline
summary(lm(pielou ~ DO_mgL, alpha)) # negative, significant - lower evenness with more oxygen
summary(lm(pielou ~ pH, alpha)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha)) 
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha)) 

alpha1 <- alpha[alpha$season != "HS",] # normoxic season
summary(lm(pielou ~ temp, alpha1)) 
summary(lm(pielou ~ sal_psu, alpha1)) # positive, significant - higher evenness with higher salinity
summary(lm(pielou ~ DO_mgL, alpha1)) 
summary(lm(pielou ~ pH, alpha1)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha1)) 
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha1)) 

alpha2 <- alpha[alpha$season == "HS",] #hypoxic season
summary(lm(pielou ~ temp, alpha2))
summary(lm(pielou ~ sal_psu, alpha2)) # marginally significant
summary(lm(pielou ~ DO_mgL, alpha2)) # marginally significant
summary(lm(pielou ~ pH, alpha2)) 
summary(lm(pielou ~ as.numeric(Chlorophyll_ugL), alpha2)) # marginally significant
summary(lm(pielou ~ as.numeric(BGA_ugL), alpha2)) # negative, significant - low evenness with more BGA
```

</details>

Plots of Evenness.

```{r, echo=FALSE, layout="l-body-outset", fig.width=8, fig.height=4}
DO.even <- ggplot(alpha_s, aes(x = DO_mgL, y = pielou, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(DO ~ (mg~L^-1))) + 
  ylab("Pielou's Evenness") +
  ylim(0.6,1) +
  theme_bw()

sal.even <- ggplot(alpha_s, aes(x = sal_psu, y = pielou, colour = site)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey90") +
  geom_point(size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(salinity ~ (psu))) + 
  ylab("") +
  ylim(0.6,1) +
  theme_bw()

chloro.even <- ggplot(alpha_s, aes(x = as.numeric(Chlorophyll_ugL), y = pielou, 
      colour = site, shape =season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey85") +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab(bquote(Chlorophyll ~ (ug~L^-1))) + 
  ylab("") +
  ylim(0.6,1) +
  theme_bw()

bga.even <- ggplot(alpha_s, aes(x = as.numeric(BGA_ugL), y = pielou, colour = site, shape =season)) +
  geom_smooth(method = "lm", colour = "black", fill = "grey85") +
  geom_point(aes(shape = factor(season)), size = 3) +
  scale_colour_manual(values = swel_col) +
  xlab("BGA") + 
  ylab("") +
  ylim(0.6,1) +
  theme_bw()
legend <- get_legend(bga.even)

plot_grid(DO.even + theme(legend.position = 'none'), 
          sal.even + theme(legend.position = 'none'), 
          chloro.even + theme(legend.position = 'none'), 
          bga.even + theme(legend.position = 'none'), 
          legend, rel_heights = c(1,1,1,1,0.5),  ncol = 5)
```

# Region Effects

First, we look at the effect of region on diversity. The outer bay (OB) includes the two normoxic sites (Cristobal and P. Caracol) and the inner bay (IB) included the two hypoxic sites.

```{r}
anova <- aov(shannon ~ Region, alpha) # H'  significant ***
summary(anova)

anova <- aov(richness ~ Region, alpha) # Richness significant  ***
summary(anova)

anova <- aov(pielou ~ Region, alpha) # Evenness significant  ***
summary(anova)
```

Boxplots showing the significant differences between different diversity metrics and region.

```{r, echo=FALSE}
H_region <- ggplot(alpha, aes(x = Region, y = shannon, fill = Region)) +
  geom_boxplot(aes(fill = Region)) +
  scale_colour_manual(values = swel_col) +
  scale_fill_manual(values = c("grey60", "grey90"), guide = FALSE) +
  ylab("Shannon's H'") + 
  xlab('') +
  theme_bw() +
  stat_compare_means(label = "p.signif")

Rich_region <- ggplot(alpha, aes(x = Region, y = richness, fill = Region)) +
  geom_boxplot(aes(fill = Region)) +
  scale_colour_manual(values = swel_col) +
  scale_fill_manual(values = c("grey60", "grey90"), guide = FALSE) +
  ylab("Species richness") + 
  xlab('Region')+
  theme_bw() +
  stat_compare_means(label = "p.signif")

Even_region <- ggplot(alpha, aes(x = Region, y = pielou, fill = Region)) +
  geom_boxplot(aes(fill = Region)) +
  # geom_point(size = 3, aes(colour = site)) + 
  scale_fill_manual(values = c("grey60", "grey90"), guide = FALSE) +
  ylab("Pielou evenness") + 
  xlab('')+
  theme_bw()+
  stat_compare_means(label = "p.signif")

plot_grid(H_region + theme(legend.position = 'none'), 
          Rich_region + theme(legend.position = 'none'), 
          Even_region + theme(legend.position = 'none'), 
          rel_heights = c(1,1,1),  ncol = 3)
```

These one-way ANOVA tests reveals a significant relationship between region (a categorical variable) and each diversity metric. Greater macrofaunal diversity and richiness was observed in the outer bay compared to the inner bay. The outer bay had less evenness than the inner bay.

# Seasonal Effects

Does seasonality impact H' diversity, richness and evenness?

```{r, eval=TRUE}
anova_int_shannon <- aov(shannon ~ season *site, alpha)
summary(anova_int_shannon) # site X season interaction exists when 0's are kept in the dataset

anova_int_rich <- aov(richness ~ season *site, alpha)
summary(anova_int_rich) # site X season interaction exists when 0's are kept in the dataset

anova__int_even <- aov(pielou ~ season *site, alpha)
summary(anova__int_even) # not significant

```

```{r, echo=FALSE}
# my_comparisons <- list( c("Pastores", "Cristobal"), c("Almirante", "Cristobal"), c("Almirante", "P_Caracol") )
ggplot(alpha, aes(x = site, y = shannon)) +
  geom_boxplot(aes(fill = season)) +
  scale_fill_manual(values = c("grey60", "grey90")) +
  ylab("Shannon's H'") + 
  xlab('site')+
  theme_bw() +
  theme(legend.position="bottom") +
  #stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  #stat_compare_means(method = "anova", label.y=3) +  # Add global p-value
  #stat_compare_means(label = "p.signif", method = "t.test",
  #                   ref.group = "Pastores")     # Pairwise comparison against reference
  stat_compare_means(label = "p.signif", aes(group = season)) # based on wilcocon

```

The only site that in which the macrofaunal diversity is affected by seasonality is P. Caracol. Surprisingly, the diversity is higher during the hypoxic season compared to the normoxic season. 

We can see this using the package *ggpubr* and comparing the means of variable 'shannon' (i.e. Shannon's H' diversity) in the different site groups formed by the grouping variable 'season.' P-values/significance tests are based on the non-parametric wilcoxon test.

```{r, echo=FALSE}
ggplot(alpha, aes(x = site, y = richness)) +
  geom_boxplot(aes(fill = season)) +
  scale_fill_manual(values = c("grey60", "grey90")) +
  ylab("Species Richness'") + 
  xlab('site')+
  theme_bw() + 
  theme(legend.position="bottom") +
  stat_compare_means(label = "p.signif", aes(group = season))
```

During the normoxic season the macrofaunal community at Pastores has higher species richness, compared to the hypoxic season. Only at Pastores is species richness significantly influenced by season.

```{r, echo=FALSE}
ggplot(alpha, aes(x = site, y = pielou)) +
  geom_boxplot(aes(fill = season)) +
  # geom_point(size = 3, aes(colour = site)) + 
  scale_fill_manual(values = c("grey60", "grey90")) +
  ylab("Species Evenness'") + 
  xlab('site')+
  theme_bw() + 
  theme(legend.position="bottom") +
  stat_compare_means(label = "p.signif", aes(group = season))
```

Yes. Diversity appears to significantly change due to season in some sites, as seen with a significant siteX season interaction (H' and richness). All the samples without any occurances were considered leaving many 'zeros' and not fitting into the assumptions of ANOVA very well.

# Temporal Effects

```{r, eval=TRUE}
anova_time_even <- aov(pielou ~ date * site * season, alpha)
summary(anova_time_even)
```

There is a significant effect of site and the interaction between date and site on Pielou's Evenness.

```{r, echo=FALSE}
p_even_time <- ggplot(alpha, aes(x = date, y = pielou, group = site, color =site)) +
  geom_point(size = 1) +
  scale_colour_manual(values = swel_col) +
  ylab("Species Evenness'") + 
  xlab('week')+
  theme_bw() + 
  theme(legend.position="bottom") + 
  geom_smooth(span = 0.2, se = FALSE) +
  scale_x_date(breaks = date_breaks("2 month"),                
               labels = date_format("%b-%y")) 
p_even_time
#p_even_time + facet_wrap(~site)
```

From the graph here we see that Almirante is the most variable, with evenness decreases later in the year. 
 
And Shannon's H' Diversity?
 
```{r, eval=TRUE}
anova_date_H <- aov(shannon ~ date * site * season, alpha)
summary(anova_date_H)
```

Date, site  and their interaction is significant. P. Caracol and Cristobal are not effected by the temporal changes. The inner bay sites, Almirante and Pastores, however, experience decline in species' diversity during the hypoxic season, as well as periodically (Feb. 2018). Note, oxygen conditions were uncharacteristically low in Feb. 2018 explaining this decline in diversity during this period. 

```{r, echo=FALSE}
p_H_time <- ggplot(alpha, aes(x = date, y = shannon, group = site, color =site)) +
  geom_point(size = 1) +
  scale_colour_manual(values = swel_col) +
  ylab("Species Diversity H'") + 
  xlab('date')+
  theme_bw() + 
  theme(legend.position="bottom") + 
  geom_smooth(span = 0.2, se = FALSE) +
  scale_x_date(breaks = date_breaks("2 month"),                
               labels = date_format("%b-%y")) 
p_H_time
#p_H_time +  facet_wrap(~site)
```

What about Richness?
 
```{r, eval=TRUE}
anova_date_rich <- aov(richness ~ date * site * season, alpha)
summary(anova_date_rich)
```

Date and site are significant. No interactions. In terms of species richness, this plot shows that there is a general pattern with the highest richness in the most northern, outer bay site (P. Caracol), followed by Cristobal. Almirante and Pastores, the two inner bay sites are similar through the year. 

All sites experience a reduction in species richness between Sept-Nov. 2017 (corresponding the the hypoxic event in the bay), as will as in February, July, November 2018. 


```{r, echo=FALSE}
p_rich_time <- ggplot(alpha, aes(x = date, y = richness, group = site, color =site)) +
  geom_point(size = 1) +
  scale_colour_manual(values = swel_col) +
  ylab("Species Richness") + 
  xlab('date')+
  theme_bw() + 
  theme(legend.position="bottom") + 
  geom_smooth(span = 0.2, se = FALSE) +
  scale_x_date(breaks = date_breaks("2 month"),                
               labels = date_format("%b-%y")) 
p_rich_time
#p_rich_time +  facet_wrap(~site)
```

```{r include=FALSE, eval=TRUE}
remove(list = ls())
```

## Source Code {.appendix}

The source code for this page can be accessed on GitHub by [clicking this link](https://github.com/tropical-repo/web/blob/master/trepo-macroalpha.Rmd).
