---
title: "Physical Conditions"
description: |
  Reproducible workflow for ... In this workflow, ....
author:
#  - name: Noelle Lucey
#    url: https://example.com/norajones
#    affiliation: Spacely Sprockets
#    affiliation_nrl: https://example.com/spacelysprokets
bibliography: assets/cite.bib
output:
    distill::distill_article:
      css: assets/styles.css
      toc: true
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(119)
library("ggplot2")
library("scales")
library("forecast")
library("reshape2")
library("xts")
library("smooth")
library("Mcomp")
library("dygraphs")
library("plotly")
library("TTR")
library("quantmod")

options(scipen=999)
knitr::opts_current$get(c(
  "cache",
  "cache.path",
  "cache.rebuild",
  "dependson",
  "autodep"
))
```

> Hit the *Hide Code* button to hide the R code (shown by default).

<aside>
```{r codefolder_ssu18, echo=FALSE, results='asis', eval=TRUE}
codefolder::generic(init = "show", query = "pre.sourceCode",
  style = "position: absolute; right: 14%; z-index: 200")
```
</aside>

```{r include=FALSE, eval=TRUE}
## Load to build page only #2
remove(list = ls())
load("page_build/trepo/physical_wf.rdata")
```

# Methods

Weekly measurements of physical conditions were collected from October 2017 to March 2020 at four seafloor sites in Almirante Bay. This was done by dangling a YSI Sonde from a boat and collecting measurements approx. 1 m from the seafloor, at 20 ± 5 m (EXO2 & EXO optical DO; Yellow Springs, USA). These sensors were calibrated monthly following the manufacturer’s instructions.

Sites are colored in the below time-series graphs for each parameter. You can **highlight** a site on the graph by pointing to it with your cursor. Specific values along the timeseries will show up on the top of the graph. You can also **zoom** into a time period of interest by moving the range selector at the bottom of the graph. 

<hr/>

**Sites**  
- Almirante: seasonal low oxygen for past 8 years  
- Pastores: seasonal low oxygen for past 8 years  
- Cristobal: normoxic for past 8 years  
- Punta Caracol: normoxic for past 8 years  

<hr/>

Sensor accuracy and precision:

```
1. Dissolved Oxygen: accuracy ± 0.1 mg/L
2. Temperature: accuracy ±0.01°C
3. pH: accuracy 0.02 units
4. Chlorophyll: add
5. Cyanobacteria: add
```

```{r, include=FALSE}
ds <- read.csv("files/trepo/physical/tables/Tropical_REPO_raw_narm.csv",
               stringsAsFactors = FALSE)

# This version no missing values (i.e. deleted), try to find and fix
# Fix the dataset to include dates from previous August so smoothing does not cut out
# the first part of the sampling period

# Deal with the date formatting so time-series analyses work:
ds$date <- as.POSIXct(ds$date, format = "%m/%d/%y", tz = "EST")
dateWindow <- c("2017-10-04", "2020-02-04")
#str(ds)
```

# Time-series Analysis

## Dissolved Oxygen

First we look at dissolved oxygen values by site over time. Here the function `na.locf`deals with any missing data and uses last value to fill gaps.  Simple Moving Average (`SMA`) is a method of time series smoothing and is actually a very basic forecasting technique. It does not need estimation of parameters, but rather is based on order selection. It is a part of smooth package. Here, `SMA` is a simple moving average that chooses 14 days but 30 days shows clear seasonal signal. At the end of this code block, a raw time-series object with all sites and a smoothed time-series object with all sites are generated.

```{r}
alm <- ds[which(ds$site == 'Almirante'),]
Alm <- xts(x = alm$DO_mgL, order.by = alm$date)
almir_narm <- na.locf(Alm, fromLast = TRUE)   
Almirante <- SMA(almir_narm, n = 30) 

past <- ds[which(ds$site == 'Pastores'),]
Past <- xts(x = past$DO_mgL, order.by = past$date)
past_narm <- na.locf(Past, fromLast = TRUE)  
Pastores <- SMA(past_narm, n = 30)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.Cara <- xts(x = pucl$DO_mgL, order.by = pucl$date)
pcara_narm <- na.locf(P.Cara, fromLast = TRUE)  
P.Caracol <- SMA(pcara_narm, n = 30)

cris <- ds[which(ds$site == 'Cristobal'),]
Cris <- xts(x = cris$DO_mgL, order.by = cris$date)
Cris <- na.locf(Cris, fromLast = TRUE)
Cristobal <- SMA(Cris, n = 30)

DO_mgL_raw <- cbind(Alm,Past,P.Cara,Cris)
DO_mgL_smoothed <- cbind(Almirante,Pastores,Cristobal, P.Caracol)
```

#### SMOOTHED time-series data 

```{r}
smoothed_do <- dygraph(DO_mgL_smoothed, main = "Dissolved Oxygen (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyAxis("y", label = "DO mg/L") %>% 
  dyAxis("x", label = "time") %>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_do
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_do
```

#### RAW time-series data 

```{r}
raw_do <- dygraph(DO_mgL_raw, main = "Dissolved Oxygen (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dySeries("Alm", label = "Almirante", color = "#CC79A7") %>%
  dySeries("Past", label = "Pastores", color = "#E69F00") %>%
  dySeries("Cris", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.Cara", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_do
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_do
```


This is the decomposed dissolved oxygen time series from the Almirante site. Again, `na.locf` deals with any missing data and uses last value to fill gaps.

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_last <- na.locf(Almirante, fromLast = TRUE) 
# plot(xts_last)
# defines time series data as weekly 
xts_ts <- ts(as.numeric(xts_last), frequency = 365.25/7) 
# str(xts_ts) # this should turn it into one column of TS
# calculates all the elements of the TS object, raw, trend, seasonal random
md <- decompose(xts_ts) 
#see decomposed time-series plots
plot(md) 
# mdadj <- xts_ts - md$seasonal # subtract seasonality from TS to reduce noise
# plot(mdadj) 
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md)
```

## Temperature

Next, we look at temperature values by site over time. The code here is repeated as above.

```{r}
alm <- ds[ which(ds$site == 'Almirante'),]
AlmT <- xts(x = alm$temp, order.by = alm$date)
almir_narmT <- na.locf(AlmT, fromLast = TRUE) 
AlmiranteT <- SMA(almir_narmT)

past <- ds[which(ds$site == 'Pastores'),]
PastT <- xts(x = past$temp, order.by = past$date)
past_narmT <- na.locf(PastT, fromLast = TRUE)  
PastoresT <- SMA(past_narmT)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.CaraT <- xts(x = pucl$temp, order.by = pucl$date)
pcara_narmT <- na.locf(P.CaraT, fromLast = TRUE)  
P.CaracolT <- SMA(pcara_narmT)

cris <- ds[which(ds$site == 'Cristobal'),]
CrisT <- xts(x = cris$temp, order.by = cris$date)
CrisT <- na.locf(CrisT, fromLast= TRUE)
CristobalT <- SMA(CrisT)

temp_raw <- cbind(AlmT,PastT,CrisT,P.CaraT)
temp_smoothed <- cbind(AlmiranteT,PastoresT,CristobalT, P.CaracolT)
```

#### SMOOTHED time-series data 

```{r}
smoothed_temp <- dygraph(temp_smoothed, main = "Temperature (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyAxis("y", label = "Temp. C") %>% 
  dyAxis("x", label = "time") %>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_temp
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_temp
```

#### RAW time-series data

```{r}
raw_temp <- dygraph(temp_raw, main = "Temperature (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dySeries("AlmT", label = "Almirante", color = "#CC79A7") %>%
  dySeries("PastT", label = "Pastores", color = "#E69F00") %>%
  dySeries("CrisT", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.CaraT", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_temp
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_temp
```

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_lastT <- na.locf(AlmiranteT, fromLast = TRUE)
# plot(xts_last)
 # defines time series data as weekly 
xts_tsT <- ts(as.numeric(xts_lastT), frequency = 365.25/7)
# str(xts_ts) # this should turn it into one column of TS
 # calculates all the elements of the TS object, raw, trend, seasonal random
mdT <- decompose(xts_tsT)
plot(mdT)
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md)
```

## pH

Next, we look at pH values by site over time.

```{r}
alm <- ds[ which(ds$site == 'Almirante'),]
Alm_pH <- xts(x = alm$pH, order.by = alm$date)
almir_narm_pH <- na.locf(Alm_pH, fromLast= TRUE)
Almirante_pH <- SMA(almir_narm_pH)

past <- ds[which(ds$site == 'Pastores'),]
Past_pH <- xts(x = past$pH, order.by = past$date)
past_narm_pH <- na.locf(Past_pH, fromLast= TRUE)  
Pastores_pH <- SMA(past_narm_pH)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.Cara_pH <- xts(x = pucl$pH, order.by = pucl$date)
pcara_narm_pH <- na.locf(P.Cara_pH, fromLast= TRUE)  
P.Caracol_pH <- SMA(pcara_narm_pH)

cris <- ds[which(ds$site == 'Cristobal'),]
Cris_pH <- xts(x = cris$pH, order.by = cris$date)
Cris_pH <- na.locf(Cris_pH, fromLast= TRUE)
Cristobal_pH <- SMA(Cris_pH)

pH_raw <- cbind(Alm_pH, Past_pH, Cris_pH, P.Cara_pH)
pH_smoothed <- cbind(Almirante_pH, Pastores_pH, Cristobal_pH, P.Caracol_pH)
```

#### SMOOTHED time-series data 

```{r}
smoothed_pH <- dygraph(pH_smoothed, main = "pH (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>% 
  dyAxis("y", label = "pH")%>% 
  dyAxis("x", label = "time")%>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_pH
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_pH
```

#### RAW time-series data

```{r}
raw_pH <- dygraph(pH_raw, main = "pH (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>% 
  dySeries("Alm_pH", label = "Almirante", color = "#CC79A7") %>%
  dySeries("Past_pH", label = "Pastores", color = "#E69F00") %>%
  dySeries("Cris_pH", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.Cara_pH", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_pH
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_pH
```

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_last_pH <- na.locf(Alm_pH, fromLast= TRUE)
# plot(xts_last)
xts_ts_pH <- ts(as.numeric(xts_last_pH), frequency = 365.25/7)

md_pH <- decompose(xts_ts_pH)
plot(md_pH)
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md_pH)
```

## Salinity

Then salinity values by site over time.

```{r, salinity plots, echo=TRUE}
alm <- ds[ which(ds$site == 'Almirante'),]
Alm_sal <- xts(x = alm$sal_psu, order.by = alm$date)
almir_narm_sal <- na.locf(Alm_sal, fromLast= TRUE)
Almirante_sal <- SMA(almir_narm_sal)

past <- ds[which(ds$site == 'Pastores'),]
Past_sal <- xts(x = past$sal_psu, order.by = past$date)
past_narm_sal <- na.locf(Past_sal, fromLast= TRUE)  
Pastores_sal <- SMA(past_narm_sal)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.Cara_sal <- xts(x = pucl$sal_psu, order.by = pucl$date)
pcara_narm_sal <- na.locf(P.Cara_sal, fromLast= TRUE)  
P.Caracol_sal <- SMA(pcara_narm_sal)

cris <- ds[which(ds$site == 'Cristobal'),]
Cris_sal <- xts(x = cris$sal_psu, order.by = cris$date)
Cris_sal <- na.locf(Cris_sal, fromLast= TRUE)
Cristobal_sal <- SMA(Cris_sal)

sal_raw <- cbind(Alm_sal,Past_sal,Cris_sal,P.Cara_sal)
sal_smoothed <- cbind(Almirante_sal,Pastores_sal,Cristobal_sal, P.Caracol_sal)
```

#### SMOOTHED time-series data 

```{r}
smoothed_sal <- dygraph(sal_smoothed, main = "Salinity (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>% 
  dyAxis("y", label = "psu")%>% 
  dyAxis("x", label = "time")%>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_sal
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_sal
```

#### RAW time-series data 

```{r}
raw_sal <- dygraph(sal_raw, main = "Salinity (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>% 
  dySeries("Alm_sal", label = "Almirante", color = "#CC79A7") %>%
  dySeries("Past_sal", label = "Pastores", color = "#E69F00") %>%
  dySeries("Cris_sal", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.Cara_sal", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_sal
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_sal
```

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_last_sal <- na.locf(Alm_sal, fromLast= TRUE)
# plot(xts_last)
xts_ts_sal <- ts(as.numeric(xts_last_sal), frequency = 365.25/7)
# str(xts_ts) # this should turn it into one column of TS
md_sal <- decompose(xts_ts_sal)
plot(md_sal)
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md_sal)
```

## Chlorophyll 

Chlorophyll values by site over time.

```{r, chlorophyll plots, echo=TRUE}
alm <- ds[ which(ds$site == 'Almirante'),]
Alm_chl <- xts(x = alm$Chlorophyll_ugL, order.by = alm$date)
almir_narm_chl <- na.locf(Alm_chl, fromLast = TRUE)
Almirante_chl <- SMA(almir_narm_chl)

past <- ds[which(ds$site == 'Pastores'),]
Past_chl <- xts(x = past$Chlorophyll_ugL, order.by = past$date)
past_narm_chl <- na.locf(Past_chl, fromLast = TRUE)  
Pastores_chl <- SMA(past_narm_chl)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.Cara_chl <- xts(x = pucl$Chlorophyll_ugL, order.by = pucl$date)
pcara_narm_chl <- na.locf(P.Cara_chl, fromLast = TRUE)  
P.Caracol_chl <- SMA(pcara_narm_chl)

cris <- ds[which(ds$site == 'Cristobal'),]
Cris_chl <- xts(x = cris$Chlorophyll_ugL, order.by = cris$date)
Cris_chl <- na.locf(Cris_chl, fromLast= TRUE)
Cristobal_chl <- SMA(Cris_chl)

chl_raw <- cbind(Alm_chl,Past_chl,Cris_chl,P.Cara_chl)
chl_smoothed <- cbind(Almirante_chl,Pastores_chl,Cristobal_chl, P.Caracol_chl)
```

#### SMOOTHED time-series data 

```{r}
smoothed_chl <- dygraph(chl_smoothed, main = "Chlorophyll (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyAxis("y", label = "ug/L") %>% 
  dyAxis("x", label = "time") %>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_chl
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_chl
```

#### RAW time-series data

```{r}
raw_chl <- dygraph(chl_raw, main = "Chlorophyll (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dySeries("Alm_chl", label = "Almirante", color = "#CC79A7") %>%
  dySeries("Past_chl", label = "Pastores", color = "#E69F00") %>%
  dySeries("Cris_chl", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.Cara_chl", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_chl
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_chl
```

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_last_chl <- na.locf(Alm_chl, fromLast = TRUE)
# plot(xts_last)
xts_ts_chl <- ts(as.numeric(xts_last_chl), frequency = 365.25/7)
# str(xts_ts) 
md_chl <- decompose(xts_ts_chl)
plot(md_chl) 
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md_chl)
```

## BGA

And finally, BGA values by site over time.

```{r}
alm <- ds[ which(ds$site == 'Almirante'),]
Alm_bga <- xts(x = alm$BGA_ugL, order.by = alm$date)
almir_narm_bga <- na.locf(Alm_bga, fromLast = TRUE)
Almirante_bga <- SMA(almir_narm_bga)

past <- ds[which(ds$site == 'Pastores'),]
Past_bga <- xts(x = past$BGA_ugL, order.by = past$date)
past_narm_bga <- na.locf(Past_bga, fromLast = TRUE)  
Pastores_bga <- SMA(past_narm_bga)

pucl <- ds[which(ds$site == 'P_Caracol'),]
P.Cara_bga <- xts(x = pucl$BGA_ugL, order.by = pucl$date)
pcara_narm_bga <- na.locf(P.Cara_bga, fromLast = TRUE)  
P.Caracol_bga <- SMA(pcara_narm_bga)

cris <- ds[which(ds$site == 'Cristobal'),]
Cris_bga <- xts(x = cris$BGA_ugL, order.by = cris$date)
Cris_bga <- na.locf(Cris_bga, fromLast= TRUE)
Cristobal_bga <- SMA(Cris_bga)

bga_raw <- cbind(Alm_bga,Past_bga,Cris_bga,P.Cara_bga)
bga_smoothed <- cbind(Almirante_bga,Pastores_bga,Cristobal_bga, P.Caracol_bga)
```

#### SMOOTHED time-series data 

```{r}
smoothed_bga <- dygraph(bga_smoothed, main = "Cyanobacteria (Smoothed)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyAxis("y", label = "ug/L") %>% 
  dyAxis("x", label = "time") %>% 
  dySeries("SMA", label = "Almirante", color = "#CC79A7") %>%
  dySeries("SMA.1", label = "Pastores", color = "#E69F00") %>%
  dySeries("SMA.2", label = "Cristobal", color = "#0072B2") %>%
  dySeries("SMA.3", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
smoothed_bga
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
smoothed_bga
```


#### RAW time-series data

```{r}
raw_bga <- dygraph(bga_raw, main = "Cyanobacteria (Raw)", width = "1000") %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dySeries("Alm_bga", label = "Almirante", color = "#CC79A7") %>%
  dySeries("Past_bga", label = "Pastores", color = "#E69F00") %>%
  dySeries("Cris_bga", label = "Cristobal", color = "#0072B2") %>%
  dySeries("P.Cara_bga", label = "P.Caracol", color = "#56B4E9") %>%
  dyRangeSelector(dateWindow = dateWindow) %>%
  dyLegend(width = 400)
raw_bga
```

<br/>

```{r, echo=FALSE, eval=TRUE, layout="l-body", fig.height=5}
raw_bga
```

#### Time-series analysis; per site for a specific water parameter 

```{r}
xts_last_bga <- na.locf(Alm_bga, fromLast= TRUE)
# plot(xts_last)
xts_ts_bga <- ts(as.numeric(xts_last_bga), frequency = 365.25/7)
# str(xts_ts) 
md_bga <- decompose(xts_ts_bga)
plot(md_bga)
```

<br/>

```{r, echo=FALSE, eval=TRUE}
plot(md_bga)
```

```{r include=FALSE}
## Load to build page only #2
save.image("page_build/trepo/physical_wf.rdata")
```

```{r include=FALSE, eval=TRUE}
## Load to build page only #2
remove(list = ls())
```


##  Source Code {.appendix}

The source code for this page can be accessed on GitHub by [clicking this link](https://github.com/tropical-repo/web/blob/master/trepo-physical.Rmd).
