---
title: "2. Macrofauna Beta diversity"
description: |
  
author:
#  - name: Noelle Lucey
#    url: https://example.com/norajones
#    affiliation: Spacely Sprockets
#    affiliation_nrl: https://example.com/spacelysprokets
# bibliography: assets/cite.bib
output:
    distill::distill_article:
     # css: assets/styles.css
      toc: true
      toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set.seed

library(vegan)
library(ggplot2)
library(viridis)
library(cowplot)
```

```{r include=FALSE, eval=TRUE}
## Load to build page # 2 trepo-macrobeta.rmd
remove(list = ls())
# load("page_build/trepo/beta_ssu_full_part_1_wf.rdata")
```

# Species Composition

In this section, we explore the differences in the macrofaunal
assemblages between the four sites across Almirante Bay weekly
throughout a one-year time frame.

1.  First we determine if there differences in the species composition
    of the marine taxa sampled? Do some taxa prefer particular sites
    while others are more generalized?

2.  Then we look at seasonal influences. Does the hypoxic season impact
    their abundance? Which species are most impacted by seasonal
    changes? Which species are not?

3.  Finally we look at the assemblage composition differences based on
    environmental parameters. This is because seasonality may not be the
    best measure of actual environmental changes. With the physical
    parameters page (add link) we see extreme fluctuatios, sometimes on
    a weekly basis, highlighting the stoicasticisty of the conditions.
    This is particularly true for dissolved oxygen.

We compare the dissimilarity in community composition between each site
using two metrics; Bray-Curtis dissimilarity and Jaccard dissimilarity.
Bray-Curtis Dissimilarity is an abundance-based dissimilarity metric
while Jaccard Dissimilarity explores the intersection over the union of
sites (i.e. turnover).

These questions are broken down in two parts. First, we statistically
answer these questions (Part 1) then we visualize the relationships with
different distance-based similarity matrices and ordination methods
(Part 2).

Ordination methods:

-   nMDS

-   PCoA

-   RDA (Redundancy analysis)

-   CAP (canonical analysis of principal coordinates)- show the effect
    of the interaction between factors on your community

-   dbRDA (distance based redundancy analysis

Similarity matrix:

-   Euclidean - good for environmental variables

-   Bray-Curtis (=Sorensen) - good for species variables

-   Jaccard

# Part 1: Statistical Analysis

```{r, include=FALSE}
swel_col <- c("#E69F00", "#CC79A7", "#0072B2", "#56B4E9")
```

```{r, include=FALSE}
ds <- read.csv("T_REPO_raw.csv", header = TRUE)
ds <- ds[ds$size == "combined",] 
ds$site <- ordered(ds$site, levels = c("Pastores", "Almirante", "Cristobal", "P_Caracol"))
ds <- ds[ds$week != "WXX",] 
site <- factor(ds$site) 
season <- factor(ds$season)
week <- factor(ds$week)
```

Look at untransformed abundance data again:

```{r}
par(mar=c(2,10,2,2)) 
boxplot(ds[,16:32],horizontal = TRUE,las=2, main="Abundance")
```

It looks like some species are much more abundant and variable than
others. It's probably a good idea to check our mean-variance
relationship!

First we convert the species data to an mvabund object format used by
the mvabund function. The function \`mvabund\`calculates the abundance
in each sample based on counts of all of the lowest taxonomic groups
identified. It is a part of mvabund package. Data is transformed
(square-root) (\`sqrt\`) to better visualize the differences and is
actually a very common transformation for abundance data.

```{r}
ds_spp <- mvabund(ds[,16:32]) 
```

Then we look at the mean-variance relationship using the meanvar.plot
function:

objects()
gdata

```{r}
meanvar.plot(ds_spp)
```

You can clearly see that the species with high means (on the x axis)
also have high variances (y axis). Then we fit predictive models using a
negative binomial model for counts.

The mvabund approach improves power across a range of species with
different variances and includes an assumption of a mean-variance
relationship use the abundance data. This is family of GLMs with an
appropriate mean-variance assumption. The default family used by mvabund
when fitting multivariate GLMs is negative binomial which assumes a
quadratic mean-variance relationship and a log-linear relationship
between the response variables and any continuous variables.

To look at the transformed abundance data we use the mvabund plot
function:

```{r}
transformed_abund <- plot(ds_spp ~ ds$site, cex.axis=0.8, cex=0.8)
```

This shows the most abundant species by site. Non-mobile polychaete
worms (Sedentaria) appear to be the most abundant of the different taxa,
and P. Caracol appears to have the highest abundances of many taxa.

## Site x Season

```{r}
model <- manyglm(ds_spp ~ site * season, family="negative_binomial")
M_output <- anova(model)
```

```{r, eval=TRUE}
M_output
```

Check our model assumptions.

```{r}
model1_res_plot <- plot(model) 
```

The residuals look good with the negative binomial distribution (i.e no
patterns detectable). 
Note: no samples with were removed from this dataset. We can remove the species with lowest abundances (i.e. Echinoidea, Scaphopoda, Amphipoda, Cladocera, Cumacea which all had low counts within the full bay-wide dataset). This would help the distribution more.. For now we keep everything. 

There were different species assemblages in the different sites (LRT =
710, P\< 0.001) and during different seasons (LRT = 89.6, P \< 0.001).
There was also an interaction between region and season (LRT = 55, P \<
0.003).

Then we perform pairwise comparisons to determine which sites are
different from each other. For the paper, increase the nBoot to 999...


```{r, eval=TRUE}
model_pairs <- anova(model, pairwise.comp = site, nBoot = 199)
model_pairs
```

We see from these results that there is a significant effect of site,
meaning that the species composition of macrofauna clearly differs
between the sites that they are found in. Every site is different from
the others, except Almirante and Pastores.

To determine which species are important, in this we run univariate
tests for each species separately. We also want to know if there is an
interaction between site and seasonality in species abundances. This is
done by using the p.uni="adjusted" argument in the anova function. The
"adjusted" part of the argument refers to the resampling method used to
compute the p values, taking into account the correlation between the
response variables. This correlation is often found in ecological
systems where different species will interact with each other, competing
with or facilitating each others' resource use.

Testing hypotheses about the site effects and site-by-season
interactions, again using a Wald statistic and 199 resamples (better to
ramp up to 999 for the paper):

```{r, eval=TRUE}
M1_uni <- anova(model, p.uni="adjusted") 
M1_uni
```

Significant Site Effects:

-   unknown, unidentified taxa (LRT = 35.635, P = 0.001)

-   Sedentaria - worms that don't move (sedentary)

-   Errantia - worms that move

-   Caudofoveata

-   Ostracoda

Significant Season Effects:

-   unknown, unidentified taxa

-   Caudofoveata (LRT = 11.5, P =0.22)

-   Bivalvia (P= 0.002)

Significant Site X Season Interaction:

-   Errantia (LRT= 20.287, P = 0.008) - worms that can move around

Errantia has a significant interaction between site and season
indicating maybe they repopulate uninhabitable sites most quickly, or be
more transient moving through the bay (LRT = 18, P = 0.002). This may be
a good indicator of the environmental conditions changing.

Check regional differences in species abundance. Testing hypotheses
about the region effect and region-by-season interactions, again using a
Wald statistic and 199 resamples (better to ramp up to 999 for the
paper):

## Site x Region

```{r, eval=TRUE}
M2 <- manyglm(ds_spp ~ ds$Region * ds$season, family="negative_binomial")
M2 <- anova(M2)
```

There were different species assemblages in the different regions (LRT =
710, P\< 0.001) and during different seasons (LRT = 89.6, P \< 0.001).
There was not a strong interaction between region and season (LRT = 27,
P \< 0.067).

Again, which species is important in this interaction?

```{r, eval=FALSE}
M2 <- manyglm(ds_spp ~ ds$Region * ds$season, family="negative_binomial")
M2_uni <- anova(M2, p.uni="adjusted") 
```

```{r, echo=TRUE}
M2_uni 
```
Significant Region Effect:

-   Unknown

-   Sedentaria

-   Errantia

-   Ostracoda

Significant Season Effect:

-   Unknown

-   Bivalve

-   Caudofoveata

As in the site x season model, regional differences are important for Sedentaria, Errantia/Aciculata, with the addition of Ostracods. All have significantly different abundances depending on the
Region). Only bivalves and caudofoveata are seasonally influenced (both molluscs; same as site x season effects).

## Site x Time
Week x Site

```{r, eval=TRUE}
model_wk <- manyglm(ds_spp ~ site * week, family="negative_binomial")
M_output_wk <- anova(model_wk)
M_output
```


# Part 2: Visualizing Results

```{r, include=FALSE}
# Full year-long data-set
ds <- read.csv("T_REPO_raw.csv", header = TRUE)


# This is the sum of occurrences at each site, in each season every other week (based on microbial sampling); environmental data is averaged by season and site
short <- read.csv("macro_short1.csv", header = TRUE)
short$site <- ordered(short$site, levels = c("Pastores", "Almirante", "Cristobal", "P_Caracol"))
short.data <- short[,13:29] # species (meio.data)
env_s.data <- short[,31:39] # temp, DO %, DO mgL, pH, chloro, bga (env.data)
site_s.data <- short[,1:3] # site, region, season (site.data)
```

## Calculate pairwaise dissimilarity

```{r bray}
meio.mdf <- as.matrix.data.frame(meio.data)
rownames(meio.mdf) <- site.data$site
meio.bray <- vegdist(meio.mdf, method = "bray")
meio.bray
```

```{r jaccard}
meio.jac <- vegdist(meio.mdf, method = "jaccard", binary = T)
meio.jac
```

Jaccard dissimilarity simple compares the overlap in species
presence/absence between pairs of communities, consequently two
communities with identical species present will have a Jaccard score of
0 regardless of the relative abundance of these organisms in the
community. Bray is based on who's who (i.e. abundances) and can be quite
different than Jaccard in terms of interpreting community changes.

While summary insights into community dissimilarity are valuable for
understanding the factors which determine the structure of communities
through time and space, it is important to consider which species are
driving these trends. We can identify the species which contribute most
to Bray-Curtis dissimilarity between each pair of sites using SIMPER
analysis.

SIMPLER is a method for identifying which species contribute most to
beta diversity (Bray-Curtis dissimilarity). SIMPER identifies the most
abundant and/or most variable species in the dataset. *Note that SIMPER
is severely biased towards highly abundant species.*

We can use the simper function in the R package vegan to calculate the
contribution of each species to Bray-Curtis dissimilarity between sites
in Almirante Bay. The simper function will automatically perform
pairwise comparisons between sites, and return cumulative contribution
scores up to a maximum of 0.7 (70% of total dissimilarity).

```{R simper}
#simper(meio.data, site.data$site, permutations = 999)
```

It appears the species that contribute to the majority of site
differences are the sedentary polychaete worms, at least between
Pastores, Almirante, Cristobal and P. Caracol. All these relationships
had low dissimilarity scores. Sipuncula (a mobile worm-like animal) is
important between Cristobal and Pastores as well as Almirante. The
absense of these creatures in Almirante and Pastores (check) likely has
a role in decreasing the bioturbation in these sites, leading to higher
site evenness but lower richness.

## Ordination

### Principal Coordinates Analysis (PCoA)

Now that we have calculated our dissimilarity matrices, we can using
ordination to visualize the structure of these communities. Ordination
by definition is the representation of large numbers of sites or species
in low dimensional space, such that any underlying relationships in
these data become visually apparent. Below we will use functions in the
R package vegan to perform and plot principal coordinates analysis for
both Bray-Curtis and Jaccard dissimilarity matrices to vizualise
macrofaunal community structure across Almirante Bay.

```{R}
# calculate principal coordinates analysis (Bray-Curtis)
pcoa.meio.bray <- cmdscale(meio.bray, k = 2, eig = T)

# extract axis positions for each site from cmdscale object and create a dataframe for plotting
pcoa.meio.bray.plotting <- as.data.frame(pcoa.meio.bray$points)
colnames(pcoa.meio.bray.plotting) <- c("axis_1", "axis_2")
pcoa.meio.bray.plotting$site <- rownames(pcoa.meio.bray.plotting)

# calculate the proportion of variance in the data which is explained by the first two PCoA axes
pcoa.meio.bray$eig[1]/(sum(pcoa.meio.bray$eig))
```

```{R}
pcoa.meio.bray$eig[2]/(sum(pcoa.meio.bray$eig))
```

```{R}
# create a PCoA plot
pcoa.meio.bray.plot <- ggplot(pcoa.meio.bray.plotting, aes(x = axis_1, y = axis_2, colour = site)) +
  geom_point(size = 3) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.8) +
  theme_bw() + 
  xlab("PCoA 1 (57.3%)") +
  ylab("PCoA 2 (17.1%)") +
  annotate(geom = 'text', label = 'Bray-Curtis', x = Inf, y = -Inf, hjust = 1.15, vjust = -1)
pcoa.meio.bray.plot

# repeat process with Jaccard dissimilarity matrix
pcoa.meio.jac <- cmdscale(meio.jac, k = 2, eig = T)

pcoa.meio.jac.plotting <- as.data.frame(pcoa.meio.jac$points)
colnames(pcoa.meio.jac.plotting) <- c("axis_1", "axis_2")
pcoa.meio.jac.plotting$site <- rownames(pcoa.meio.jac.plotting)

pcoa.meio.jac$eig[1]/(sum(pcoa.meio.jac$eig))
pcoa.meio.jac$eig[2]/(sum(pcoa.meio.jac$eig))

pcoa.meio.jac.plot <- ggplot(pcoa.meio.jac.plotting, aes(x = axis_1, y = axis_2, colour = site)) +
  geom_point(size = 3) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.8) +
  theme_bw() + 
  xlab("PCoA 1 (60.8%)") +
  ylab("PCoA 2 (24.9%)") +
  annotate(geom = 'text', label = 'Jaccard', x = Inf, y = -Inf, hjust = 1.215, vjust = -1)

# extract plot legend
legend <- get_legend(pcoa.meio.jac.plot)

# plot Bray-Curtis PCoA and Jaccard PCoA side by side
plot_grid(pcoa.meio.bray.plot + theme(legend.position = 'none'), pcoa.meio.jac.plot + theme(legend.position = 'none'), legend, ncol = 3, rel_widths = c(1,1,0.5))
```

## Permanova

PERMANOVA (permutational multivariate analysis of variance; Anderson
2001) is non-parametric multivariate statistical test used to quantify
the impact of both continuous and categorical variables on dissimilarity
between communities. While it is valid to construct a PERMANOVA model
which includes continuous variables, in this instance we will use a
simple PERMANOVA model to test the effect of season type on macrofaunal
community composition. The input for PERMANOVA is a dissimilarity matrix
(Bray-Curtis dissimilarity in this case), and corresponding
environmental data. A resultant p-value \< 0.05 indicates that centroid
position and/or dispersion differs between the groups in the model.

As PERMANOVA is affected by both centroid position and disperison, we
perform a homogeneity of dispersions analysis using betadisper to
establish whether dispersion is homogeneous between groups (in this
case, hypoxic season (HS) and normoxic season (NS) associated
communities). We then perform PERMANOVA analysis using the adonis
function in the R package vegan.

```{R}
# Homogeneity of dispersion test
#permutest(betadisper(meio.bray, site.data$region))
```

```{R}
#adonis(meio.bray ~ season, data = site.data, permutations = 999)
```

The betadisper analysis shows that there is no signficiant difference in
dispersion between hypoxic and normoxic seasons. The PERMANOVA analysis
show that there is no significant difference in centroid position or
dispersion between seasons.

## Distance-Based Redundancy Analysis (dbRDA)

Distance-Based Redundancy Analysis (dbRDA; Legendre and Anderson, 1999)
is an extension of Redundancy Analysis (RDA) which allows the use of
non-euclidean distance matrices as inputs (e.g. Bray-Curtis
dissimilarity). The method works by first calculating a PCoA from the
dissimilarity matrix, and then subjecting the PCoA eigenvalues (which
represent dissimilarities in euclidean space) to RDA. The method aims to
detect linear relationships between environmental variables and these
dissimilarities.

dbRDA differs from PCoA in that it is a constrained analysis. While PCoA
axes are generated to explain maximum variation in the distance matrix,
dbRDA canonical axes are constructed as linear combinations of
environmental variables. Consequently these axes are constrained to the
environmental variables in the model, and the ordination will be
distinct from the PCoA. dbRDA allows us to visualize how environmental
variables constrain variation in community composition between our
sites.

As dbRDA considers multiple environmental variables which are measured
using different techniques, it is important to normalize environmental
variables such that they can be compared concurrently. This is achieved
using a z-score transformation. Moreover, dbRDA is sensitive to
multicolinearity (i.e. high correlation between environmental
variables). We can test whether any of our environmental variables are
correlated by calculating variance inflation factors (VIFs) and then
removing selected terms from the model until all VIF scores are less
than 10.

Z-Score Transformation - normalizes environmental variables by their
standard deviation from the mean

# Z-Score transform the environmental data

```{R}
# Z-Score transform your environmental data
env.data.z <- env.data
env.data.z
env.data.z$temp <- (env.data.z$temp - mean(env.data.z$temp))/sd(env.data.z$temp)
env.data.z$sal_psu <- (env.data.z$sal_psu - mean(env.data.z$sal_psu))/sd(env.data.z$sal_psu)
env.data.z$DO_mgL <- (env.data.z$DO_mgL - mean(env.data.z$DO_mgL))/sd(env.data.z$DO_mgL)
env.data.z$pH <- (env.data.z$pH - mean(env.data.z$pH))/sd(env.data.z$pH)
env.data.z$Chlorophyll_ugL <- (env.data.z$Chlorophyll_ugL - mean(env.data.z$Chlorophyll_ugL))/sd(env.data.z$Chlorophyll_ugL)
env.data.z$BGA_ugL <- (env.data.z$BGA_ugL - mean(env.data.z$BGA_ugL))/sd(env.data.z$BGA_ugL)
```

# Perform dbRDA

```{R}
# construct full model and calculate VIF, need to remove terms from the model until all VIF scores are < 10.
dbRDA.full <- capscale(meio.bray ~ temp+sal_psu+DO_mgL+pH+Chlorophyll_ugL+BGA_ugL,
                       env.data.z)
vif.cca(dbRDA.full)
```

```{R}
# construct reduced model and calculate VIF
dbRDA.mat <- capscale(meio.bray ~  temp+DO_mgL+BGA_ugL,
                      env.data.z)
vif.cca(dbRDA.mat)
```

```{R}
# test overall significance of the analysis
anova(dbRDA.mat)
```

```{R}
# test significance of each environmental variable
anova(dbRDA.mat, by = "terms")
```

```{R}
# summary of dbRDA model to extract total variance constrained and axis scores
summary(dbRDA.mat)
```

## Plot dbRDA

```{R}
smry <- summary(dbRDA.mat)
scrs <- scores(dbRDA.mat)
df1  <- data.frame(smry$sites[,1:2]) # site scores for RDA1 and RDA2
df1$site <- rownames(df1)  #add site names
df2  <- data.frame(smry$biplot[,1:2])  # mapping environmental variables

rda.plot <- ggplot(df1, aes(x=CAP1, y=CAP2, colour = site)) + 
  geom_segment(data=df2, aes(x=0, xend=CAP1, y=0, yend=CAP2), 
               color="grey50", arrow=arrow(length=unit(0.01,"npc"))) +
  geom_text(data=df2, aes(x=CAP1,y=CAP2,label=rownames(df2),
                          hjust=0.5*(1-sign(CAP1)),vjust=0.5*(1-sign(CAP2))), 
            color="grey50", size=3) +
  geom_point(size = 3) +
  scale_colour_viridis_d(option = "magma", begin = 0.2, end = 0.8) +
  geom_text(aes(label=rownames(df1),
                hjust=0,vjust=1.5), colour = "black",size=3) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_vline(xintercept=0, linetype="dotted") +
  xlim(-1.05, 1.65) +
  ylim(-1.5, 1.05) +
  xlab("RDA1 (55.2%)") + # this percentage comes from the CAP1 'importance of components:' proportion explained, which can be found in summary(dbRDA.mat) 
  ylab("RDA2 (12.3%)") + # this percentage comes from the CAP2 'importance of components:' proportion explained, which can be found in summary(dbRDA.mat) 
  coord_fixed() +
  theme_bw() 
rda.plot
```

It is critical to remember that we have reduced our model due to
colinearity between pH, salinity, and chlorophyll, but that these
factors are confounded. The relationship between constrained variance
and unconstrained variance informs us of how well the beta-diversity of
Almirante Bay can be explained by the environmental factors in our
model. While the measured environmental parameters, namely Dissolved
Oxygen concentrations, were able to explain trends in Shannon's H'
observed throughout the bay, the environmental parameters were unable to
explain a significant proportion of variance in macrofauna community
composition (beta-diversity) throughout the region.

```{R, PCA}
# Adding environmental data
# envfit()` fits vectors or planes while `ordisurf()` fits smooth, potentially non-linear surfaces
# vector fitting using pca 

pca <- rda(decostand(ds.data, method = "hellinger"), scale =TRUE) # hellinger, 
pca
biplot(pca, scaling = "species")

ev <- envfit(pca ~ ., data = env.data,
             choices = 1:2,
             scaling = "species",
             na.rm =TRUE,
             permutations = 1000)
ev
# pH and DO significant, salinity to a lesser extent, BGA marginal, temp is not.

plot(pca, display = "species", type = "n",
     scaling = "species")
points(pca, display = "species", pch= 19,
       scaling = "species")
plot(ev)
plot(ev, add = FALSE)

###
surf <- ordisurf(pca ~ DO_mgL,
                 data = env.data,
                 knots = 10,
                 pch = 16,
                 isotropic = TRUE,
                 main = NULL)
ordipointlabel(pca,
               display = "species",
               scaling = "species",
               add = TRUE)

summary(surf)
```

```{R, eclipse}
## Draws ellipsoid hulls & standard error & deviation ellipses
disp <- "sites"
scl <- "symmetric"
plot(pca, type = "n", scaling = scl,
     display = disp)
## ellipsoid hull
ordiellipse(pca, groups = site.data$site,
            kind = "ehull", col = col_vec,
            scaling = scl, lwd = 2)
## standard error of centroid  ellipse
ordiellipse(pca, groups = site.data$site,
            draw = "polygon", col = col_vec,
            scaling = scl, lwd = 2,
            label = TRUE)
```


```{r include=FALSE, eval=TRUE}
remove(list = ls())
```

*Heavily borrowed code and text for this page from:
<https://www.rpubs.com/roalle/mres_2019> - thank you!*
